---
title: "Introduction to Distributional Cost-Effectiveness Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Distributional Cost-Effectiveness Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(dceasimR)
```

## Why standard CEA ignores equity

Standard cost-effectiveness analysis (CEA) answers one question: does this
intervention generate more health per pound spent than the alternatives? It
aggregates health across all patients as if a QALY gained by the most
deprived person equals a QALY gained by the least deprived.

This approach is silent on *who* benefits — and therefore on how interventions
affect health inequalities between socioeconomic groups.

## The DCEA framework

Distributional Cost-Effectiveness Analysis (DCEA), developed by Cookson,
Griffin, Norheim and Culyer (2020), extends standard CEA by:

1. **Distributing** aggregate health gains across socioeconomic groups.
2. **Measuring** the impact on health inequality (SII, Atkinson index, etc.).
3. **Evaluating** the social welfare gain using inequality-aversion weights.
4. **Visualising** the equity-efficiency trade-off on an impact plane.

## When to use aggregate vs full-form DCEA

| Method | When to use | Data required |
|--------|-------------|---------------|
| **Aggregate DCEA** | Standard TA; disease-level HES data available | ICER, incremental QALY/cost, disease ICD |
| **Full-form DCEA** | Subgroup trial data available; HST or exceptional case | Per-group QALY/cost estimates |

NICE (2025) recommends aggregate DCEA as the default supplementary analysis
for technology appraisals where equity is relevant.

## Quick start

```{r quick-start}
result <- run_aggregate_dcea(
  icer            = 28000,
  inc_qaly        = 0.45,
  inc_cost        = 12600,
  population_size = 12000,
  wtp             = 20000,
  opportunity_cost_threshold = 13000
)

summary(result)
```

```{r impact-plane, fig.width = 6, fig.height = 5}
plot_equity_impact_plane(result)
```

## Key references

- Cookson R, Griffin S, Norheim OF, Culyer AJ (2020). *Distributional
  Cost-Effectiveness Analysis*. Oxford University Press.
  <https://doi.org/10.1093/oso/9780198838197.001.0001>

- NICE (2025). *Technology Evaluation Methods: Health Inequalities* (PMG36).

- Love-Koh J et al. (2019). Value in Health 22(5): 518-526.
